
\input{../header}

\ifpdf%
  \ifLuaTeX%
    % for lualatex
    \pdfvariable minorversion=7% chktex 1
    % omit CreationDate and ModDate keys.
    \pdfvariable suppressoptionalinfo 767% chktex 1
    % no adding to the trailer dictionary.
    \pdfvariable trailerid{}% chktex 1
    \pdfvariable suppressoptionalinfo -1% chktex 1
  \else
    \ifXeTeX%
      % for xelatex
      \special{pdf:minorversion 7}
      % TBD: find a way to express pdfinfoomitdate: necessary? 
      \special{pdf:trailerid []}
    \else
      \ifPDFTeX%
        \pdfminorversion=7         % for pdflatex
        % omit CreationDate and ModDate keys.
        % not before pdfTeX 3.14159265-2.6-1.40.17
        \pdfinfoomitdate=1                   % for pdflatex
        % no adding to the trailer dictionary.
        %\pdftrailer=0                        % for pdflatex
        \pdftrailerid{}                       % for pdflatex
        \pdfsuppressptexinfo=-1               % for pdflatex
      \else
        % Here, the tex processor is unknown. 
      \fi%pdftex
    \fi%xetex
  \fi%luatex

  \hypersetup{
    pdfinfo={
      Author      ={Ernst Reissner},
      Title       ={Presentation with/of the latex-maven-plugin },
      CreationDate={unknown},
      ModDate     ={unknown},
      Producer    ={unknown},
      Subject     ={presentations with beamer},
      Keywords    ={LaTeX;beamer}
    }
  }
\else
\fi%ifpdf

% TBD: clarify what to do if no pdf is created. 
% As soon as this question comes up in the course of new development, 
% this will be compiled as html by accident. 
% The we can see, how well this works. 

\usetheme{Berlin}
\title{Presentation with/of the \texttt{latex-maven-plugin}}
\author{E. Reissner}
\date{ernst.reissner@simuline.eu}


\hypersetup{colorlinks,linkcolor=,urlcolor=blue,citecolor=yellow}%
%\hypersetup{frenchlinks}
\begin{document}
 
\mode<article>{\maketitle}




\begin{frame}
  \titlepage%
\end{frame}

\section*{Outline}

\begin{frame}
  \tableofcontents
\end{frame}

\section{Introduction and Purpose}

This is some additional text 

\begin{frame}
  \frametitle{Purpose of these documents }
  The purpose of this document is twofold:
  %
  \begin{itemize}
    \item Give an overview over the plugin. 
    \alert{But: }
    %
    \begin{itemize}
      \item The single official description of the plugin is the manual~\cite{LatexPlugin}.
      \item The 
      \href{http://simuline.eu/LatexMavenPlugin/index.html}{project site} 
      gives already an overview. 
      \item 
      The content of this presentation is updated only by need. 
    \end{itemize}
    
    \item Demonstrate that the plugin can compile a presentation using the \texttt{beamer} class 
    and a handout written as \texttt{beamerarticle}. 

    \begin{itemize}
      \item In fact, when citing this presentation as~\cite{PresBeamer}, 
      this refers to both. 
      \item The presentation is about the plugin 
      \item The handout adds information on how to write a beamer presentation. 
      \item Both documents turn \texttt{beamer} class and package \texttt{beamerarticle} into preferred usage 
      ensuring tests. 
    \end{itemize}
  \end{itemize}
  
\end{frame}


\section{Features}

\subsection{Realized Features}

\begin{frame}
  \frametitle{Ant and Maven}
  Automatically creates documents from LaTeX sources during the Maven \texttt{site} phase 
  and in an ant run. 

  This comprises running basic tools like \texttt{lualatex} and \texttt{bibtex} 
  and rerunning them by need. 

\end{frame}

\begin{frame}
  \frametitle{Supported IO}
  Supports 
  \begin{itemize}
    \item
    many output formats like PDF, DVI, HTML, DOCX, RTF, TXT and others
    \item
    many graphical input formats like PNG, MP, FIG, gnuplot; 
    also provides a separate goal creating them, \texttt{grp} 
    \item
    bibliography, index, glossary and embedded code; in particular split index
  \end{itemize}
\end{frame}


\begin{frame}
  \frametitle{Checks}
  Check 
  %
  \begin{itemize}
    \item
    sources with \texttt{chktex} and logs the results in target and goal \texttt{chk} 
    \item
    versions of used tools via goal \texttt{vrs} 
    \item
    log files detecting errors and warnings 
    \item
    whether a document could have been reproduced, by demand 
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Orchestration and document development}
  Orchestration of various tools detecting need for execution e.g.
  of \texttt{bibtex} including \texttt{rerunfilecheck} for \texttt{lualatex} and friends.

  Supports document development, mainly by cooperating with editor, viewer and with other tools in the build chain.
  \begin{itemize}
    \item Offers installation script for extensions of VS Code.
    \item Offers configuration file \texttt{.chktexrc} for \texttt{chktex}.
    \item Can create configuration file \texttt{.latexmkrc} 
    for \texttt{latexmk} synchronized with the configuration.
    \item Offers a common header file \texttt{header.tex} 
    to unify packages loaded by latex main files.
  \end{itemize}
\end{frame}


\subsection{Planned Features}

\begin{frame}
  \frametitle{Planned Features}
  \begin{itemize}
  \item Support \texttt{biber} replacing \texttt{bibtex} as preferred tool
  \item Support \texttt{xindy} replacing \texttt{makeindex} as preferred tool
  \item Support \texttt{bib2gls} replacing \texttt{makeglossary} as preferred tool
  \item Execute \texttt{glosstex} if needed
  \item Usage of the \texttt{multibib} macros
  \item \dots
  \end{itemize}
\end{frame}

% \section{Installation}

\section{Goals}

\begin{frame}
  \frametitle{Goals}

  An overview is on the 
  \href{http://simuline.eu/LatexMavenPlugin/plugin-info.html}{goals page}. 
  
  \setbeamercolor{alerted text}{fg=blue}
  Besides goals referring to creating output of a given type 
  like \alert{\texttt{pdf}}, \alert{\texttt{odt}} and \alert{\texttt{html}}, 
  there is a goal \alert{\texttt{cfg}}, 
  which allows \setbeamercolor{alerted text}{fg=red}
  \alert con\alert fi\alert guring various output formats as \texttt{targets}. % chktex 1
  
  \setbeamercolor{alerted text}{fg=blue}
  One further goal, \alert{\texttt{chk}} to perform a check with \texttt{chktex} 
  is allowed among the \texttt{targets}. 

  The other goals cannot serve as targets: 
  %
  \begin{description}
    \item[\texttt{vrs}] to check versions of tools, 
    \item[\texttt{grp}] to create graphic files speeding up use of \texttt{latexmk} 
    and 
    \item[\texttt{inj}] to inject useful files like headers or config files 
    like \texttt{.latexmkrc}. 
  \end{description}
  

\end{frame}

% \section{Usage}

% \section{Examples}

\mode<presentation>
\section{References}% if with star not in toc 


\begin{frame}[allowframebreaks]
  \frametitle{References}
  \bibliographystyle{alpha}
  \bibliography{../lit}{}
\end{frame}

\mode<article>
\bibliographystyle{alpha}
\bibliography{../lit}{}

\end{document}